---
title: "Background notes"
author: "David"
date: "02/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GIS plotting

Using standard packages

```{r}
library(sp)
library(spData)
library(sf)
library(tidyverse)
```

What countries do we have?

```{r}
a=world %>% 
  select(region_un,iso_a2,name_long) %>%
  filter(region_un == "Europe") 
a$name_long
```

Can we plot them? Gives axes ...

```{r}
plot(subset(world,region_un=="Europe")["name_long"])
axis(1)
axis(2)
```

but the cropping is a little haphazard

```{r}
plot(subset(world,region_un=="Europe")["name_long"],
     xlim=c(0,0.5))
axis(1)
axis(2)
```


Is ggplot better? See https://r-spatial.github.io/sf/articles/sf5.html for details


```{r}
ggplot()+geom_sf(data=a,aes(fill=name_long))
```

or better

```{r}
a$isSE=with(a,ifelse(name_long=="Sweden",T,F))
ggplot()+
  scale_fill_manual(values=c("grey","red")) +
  geom_sf(data=world,fill="white") + 
  geom_sf(data=a,aes(fill=isSE)) + 
  xlim(c(-10,50)) + ylim(c(35,70))
```
which looks better.

## Data

### Countries

Going to the WHO/EURO API
https://gateway.euro.who.int/en/api/specification/#nav-example-2

```{r}
library(httr)
library(jsonlite)
```

First try for a list of countries

```{r}
host="https://dw.euro.who.int"
a=GET(paste0(host,"/api/v3/countries"))
class(a)
```
```{r}
a1=httr::content(a)
```

which is a list of countries, each with a sublist of singletons for the column entries, so

```{r}
a2=plyr::ldply(a1,unlist)
head(a2)
```

### Datasets

Can get indicators by browsing from https://gateway.euro.who.int/en/themes/ and look at the head of the .csv file

```{r}
getData=function(code){
  a=GET(paste0(host,
               sprintf("/api/v3/measures/%s?output=data",code)))
  a1=httr::content(a)
  plyr::ldply(a1$data,unlist) 
}
```


```{r}
# https://gateway.euro.who.int/en/indicators/hfa_426-3050-pure-alcohol-consumption-litres-per-capita-age-15plus/
alcohol=getData("HFA_426")
head(alcohol)
str(alcohol)
```


```{r}
expect1=getData("HFA_46")
head(expect1)
```

```{r}
prop65=getData("HFA_13")
head(prop65)
```

Need to now tidy the data

```{r}

tidyData=function(w){
  w %>%
    rename(iso3=dimensions.COUNTRY,value=value.numeric) %>%
    select(fact_id,iso3,value) %>% 
    mutate(value=as.numeric(as.character(value))) %>%
    filter(iso3 != "")
}
head(tidyData(expect1))
```

which works: data looks almost normal ...

```{r}
qqnorm(tidyData(expect1)$value)
```


### Joining data

Need to merge into the GIS data

```{r}
head(world)
```

which uses iso_a2 in most places

```{r}
world[is.na(world$iso_a2),]
```

but not quite:

```{r}
subset(world,name_long == "France")
```
so will need to recode for iso_a2. Do this here, renaming the country code to iso2 to merge into the health data

```{r}
recodeWorld=function(w){
  
  # recode world data
  w[grep("France",w$name_long),"iso_a2"]="FR"
  w[grep("Norway",w$name_long),"iso_a2"]="NO"
  plyr::rename(w,c("iso_a2"="iso2"))
}
head(recodeWorld(world))
```

 ### encapuslation
 
Now coded all as functions as a library:

```{r}
library(EuData)
egData=addGisData(getDatasets())
with(egData,table(code,sex))
save(egData,file="EuData/data/egData.rda")
```




